Boosting Higgs Pair Production in the $bb\bar{b}\bar{b}$ Channel at the LHC with Multivariate Techniques

### 2015-06-04 paper/sec-mva.tex
work on MVA section

After the traditional cut-based analysis presented in the previous section,
and combining the significance of the different event categories,
we end up with a $S/\sqrt{B}$ value of AAA.
%
Taking into account our more realistic simulation of the QCD background as
compared to previous studies, this is not too bad.
%
However, the attentive reader might have noticed that we have used
relatively 
loose kinematics cuts, for example in the Higgs mass window.
%
The motivation for this is that the $n$-tuples of signal and background
events, after our basic selection, are processed though a multivariate
analysis with the motivation to increase the signal over background
significance.

We now apply a basic MVA to the final output of the resolved and boosted analyses. The MVA used is a feed-forward artificial neural network with $N_{\mathrm{par}}\times5\times3\times1$ architecture, where $N_{\mathrm{par}}$ is the number of input parameters for the network. The MVA was trained for 50,000 GA generations with a cross-entropy error function. In Figure~\ref{fig:nnresponse} the neural network classification of events is demonstrated for both the resolved and boosted cases. In Figure~\ref{fig:exampleroc} the ROC curves for the two analyses are shown, demonstrating that the neural network is able to
perform well in both cases. Figure~\ref{fig:nnweights} shows the distribution of NN weights in both cases.

>>>>

After the traditional cut-based analysis presented in the previous section,
and combining the significance of the different event categories,
we end up with a $S/\sqrt{B}$ value of AAA.
%
Taking into account our more realistic simulation of the QCD background as
compared to previous studies, this is not too bad.
%
However, the attentive reader might have noticed that we have used
relatively 
loose kinematics cuts, for example in the Higgs mass window.
%
The motivation for this is that the $n$-tuples of signal and background
events, after our basic selection, are processed though a multivariate
analysis with the motivation to increase the signal over background
significance.
%
Multivariate techniques are by now a common tool to enhance signal
significance in HEP analysis.

\subsection{MVA strategy}

The final step of our analysis is to process our three exclusive
categories of events (boosted, intermediate and resolved) though
a Multivariate Analysis.
%
The particular tool that we adopt is
a feed-forward artificial neural network.\footnote{This is the
  same type of neural networks as those used in the
  NNPDF analysis of parton distributions~\cite{Ball:2010de}.}
%
The architecture of the neural network is
\be
N_{\mathrm{var}}\times5\times3\times1 \, ,
\ee
with $N_{\mathrm{par}}$ the number of input variables for the MVA.
%
Therefore, we have two hidden layers and a single output variable, which
is the value of the discriminant to separate signal and background.\footnote{Note that in the MVA literature artificial neural networks with several hidden
  layers are known as {\it deep neural networks}.}
%
Below we list the variables that are given as input to the MVA in each
different category.
%
The MVA has trained using a Genetic Algorithm during 50K iterations,
and the figure of merit to be minimized is the cross-entropy function.
%
This figure of merit is defined as
\be
\bf DEF
\ee

For both signal and background events, the variables that we input
to the MVA are the following.
%
For the boosted category we have
\begin{itemize}
\item The invariant mass of the two mass-drop-tagged jets, $m_{\rm jet}$
\item The transverse momentum of the reconstructed Higgs pair, $p_{T,hh}$
\item The rapidity of the two mass-drop-tagged jets, $y_{\rm jet}$
\item The invariant mass of the leading Higgs candidate, $m_{h,1}$
\item The invariant mass of the sub-leading Higgs candidate, $m_{h,2}$
\item The splitting scale of the first Higgs
\item The splitting scale of the second Higgs
\item The subjetiness variable $\tau_{21,h1}$ of the leading Higgs candidate
\item The subjetiness variable $\tau_{21,h2}$
  of the sub-leading Higgs candidate
 \item The angular separation between the two tagged jets, $\Delta R_{1,2}$ 
\end{itemize}
For completeness, we give here the definition of the various substructure variables.
%
The splitting scale is defined as
\be
\bf DEF
\ee
while the $\tau_{1,2}$ subjetiness variable is defined as
\be
\bf DEF
\ee
The other variables are more conventional.

For the resolved analysis instead, the $N_{\rm var}$ kinematic variables
that are given as input to the MVA are the following
\begin{itemize}

  \end{itemize}

\subsection{Enhancing signal discrimination with the MVA}

In Figure~\ref{fig:nnresponse} the neural network classification of events is demonstrated for both the resolved and boosted cases. In Figure~\ref{fig:exampleroc} the ROC curves for the two analyses are shown, demonstrating that the neural network is able to
perform well in both cases. Figure~\ref{fig:nnweights} shows the distribution of NN weights in both cases.
